---
title: "Intermediate R course DDEA"
author: "Kasper Mortensen"
format: html
---

## Header 1

```{r setup}
library(tidyverse)
library(here)
library(snakecase)
library(fs)

source(here("R/functions.R"))

# {r setup} tells R to run this chunk first
```

## Importing new data

```{r}
here("data-raw/nurses-stress/stress/5C/5C_1586886626/HR.csv.gz") |>
  read_csv(
    show_col_types = FALSE,
    name_repair = to_snake_case, # does the same as clean_names, needs to library(snakecase)
    n_max = 100
  )

## NOTES ##
# snakecase -> all lower letters separated by _ (best practice)
# camelCase
# PascalCase
```

## Exercise: Reading other files.
```{r}
here("data-raw/nurses-stress/stress/5C/5C_1586886712/HR.csv.gz") |>
  read_csv(
    show_col_types = FALSE,
    name_repair = to_snake_case # does the same as clean_names from janitor
  )

here("data-raw/nurses-stress/stress/5C/5C_1586886712/HR.csv.gz") |>
  read_csv(
    show_col_types = FALSE,
    name_repair = to_snake_case,
    n_max = 100
  )

here("data-raw/nurses-stress/stress/6B/6B_1587569373/EDA.csv.gz") |>
  read_csv(
    show_col_types = FALSE,
    name_repair = to_snake_case
  )

here("data-raw/nurses-stress/stress/6B/6B_1587569373/EDA.csv.gz") |>
  read_csv(
    show_col_types = FALSE,
    name_repair = to_snake_case,
    n_max = 100
  )
```

## Making a function to add numbers

```{r}
# Roxygen skeleton (choose from palette with cursor inside the function)
# -> @export and @examples is for packages so they are deleted below
# A way of describing what the function does

#' Add two numbers together.
#'
#' @param num1 A number here.
#' @param num2 A number here.
#'
#' @returns Returns the sum of the two numbers.
#'
add_numbers <- function(num1, num2) {
  added <- num1 + num2
  return(added)
}

add_numbers(1, 2)
```

## Import 5C's HR data with a function
```{r}
#| eval: false
# When making names, check in console ?name if the name exists

#' Read in one nurses' stress data file.
#'
#' @param file_path Path to the data file.
#'
#' @returns Outputs a data frame/tibble.
#'
read <- function(file_path) {
  data <- file_path |>
    readr::read_csv( # readr:: added at the end of day 1
      show_col_types = FALSE,
      name_repair = snakecase::to_snake_case,
      n_max = 100
    )
  return(data)
}

# don't load packages inside functions
```

```{r}
here("data-raw/nurses-stress/stress/5C/5C_1586886626/HR.csv.gz") |>
  read()
here("data-raw/nurses-stress/stress/5C/5C_1586886712/HR.csv.gz") |>
  read()
```

## Add max rows to the function
```{r}
#| eval: false
# When making names, check in console ?name if the name exists

#' Read in one nurses' stress data file.
#'
#' @param file_path Path to the data file.
#' @param max_rows max number of rows loaded
#'
#' @returns Outputs a data frame/tibble.
#'
read <- function(file_path, max_rows = 100) {
  data <- file_path |>
    readr::read_csv(
      show_col_types = FALSE,
      name_repair = snakecase::to_snake_case,
      n_max = max_rows
    )
  return(data)
}

# Outputs only 100 rows
here("data-raw/nurses-stress/stress/5C/5C_1586886712/HR.csv.gz") |>
  read()

# Outputs only 1000 rows
here("data-raw/nurses-stress/stress/5C/5C_1586886712/HR.csv.gz") |>
  read(max_rows = 1000)

# Outputs all rows (`Inf` means infinity)
here("data-raw/nurses-stress/stress/5C/5C_1586886712/HR.csv.gz") |>
  read(max_rows = Inf)
```


```{r}
## Play around a bit
#| eval: false

# When making names, check in console ?name if the name exists

#' Read in one nurses' stress data file.
#'
#' @param file_path Path to the data file.
#' @param max_rows max number of rows loaded
#'
#' @returns Outputs a data frame/tibble.
#'
read <- function(file_path, max_rows = 100) {
  data <- read_csv(file_path,
    show_col_types = FALSE,
    name_repair = to_snake_case,
    n_max = max_rows
  )
  return(data)
}
# above works as well, and does the same according to the instructor. The other way
# with piping might be better, especially when you copy and paste function
```

## Day 2 - Using map
```{r}
#| eval: false

hr_files <- here("data-raw/nurses-stress/") |>
  fs::dir_ls(regexp = "HR.csv.gz", recurse = TRUE) # recurse -> If TRUE recurse fully, if a positive number the number of levels to recurse

hr_files

hr_data <- hr_files |>
  purrr::map(read) |> # remember not to use ()
  purrr::list_rbind(names_to = "file_path_id") # names_to take filepath

# hr_data is a list, others map_lgl, map_dbl and so on return other stuff

# hr_data[[1]] # look at a specific element using [[]] when the element is a list

# hr_data # shows all data but too heavy

#' Read all data based on filename
#'
#' @param filename tyoe of data
#'
#' @returns Returns a dataframe including all data of the specific type
read_all <- function(filename) {
  files <- here("data-raw/nurses-stress/") |>
    fs::dir_ls(regexp = filename, recurse = TRUE)
  data <- files |>
    purrr::map(read) |>
    purrr::list_rbind(names_to = "file_path_id")
  return(data)
}

hr_data <- read_all("HR.csv.gz")
eda_data <- read_all("EDA.csv.gz")
bvp_data <- read_all("BVP.csv.gz")
```

# inspecting data
```{r}
hr_data <- read_all("HR.csv.gz")

survey_data <- here("data-raw/nurses-stress/survey-results.csv.gz") |>
  read()

hr_data
survey_data
```

## extracting id
```{r}
hr_data |>
  dplyr::mutate(
    id = stringr::str_extract(
      file_path_id,
      pattern = "/stress/[:alnum:]{2}/" # {2} -> find exactly 2 of the thing we asked for
    ) |>
      str_remove("/stress/") |>
      str_remove("/"),
    .before = file_path_id
  ) |>
  dplyr::select(-file_path_id)
```

## get id as a function
```{r}
#| eval: false

#' Get participant ID
#'
#' @param data The data we want to extract id from
#'
#' @returns new dataframe with a id column and no file_path_id

get_participant_id <- function(data) {
  data <- data |>
    dplyr::mutate(
      id = stringr::str_extract(
        file_path_id,
        pattern = "/stress/[:alnum:]{2}/" # {2} -> find exactly 2 of the thing we asked for
      ) |>
        stringr::str_remove("/stress/") |>
        stringr::str_remove("/"),
      .before = file_path_id
    ) |>
    dplyr::select(-file_path_id)
  return(data)
}

# extra exercise using lookarounds in regex -> able to remove str_remove

#' Get participant ID
#'
#' @param data The data we want to extract id from
#'
#' @returns new dataframe with a id column and no file_path_id

get_participant_id <- function(data) {
  data <- data |>
    dplyr::mutate(
      id = stringr::str_extract(
        file_path_id,
        pattern = "(?<=/stress/)[:alnum:]{2}(?=/)" # {2} -> find exactly 2 of the thing we asked for
      ),
      .before = file_path_id
    ) |>
    dplyr::select(-file_path_id)
  return(data)
}

bvp_data <- read_all("BVP.csv.gz") |>
  get_participant_id()

bvp_data
```

## Working with across
```{r}
hr_data <- read_all("HR.csv.gz") |>
  get_participant_id()

hr_data

hr_data |>
  summarise(across(hr, mean))

hr_data |>
  summarise(
    across(
      hr,
      list(mean = mean, median = median, sd = sd) # with 'list' we can give it names
    )
  )

hr_data |>
  summarise(
    across(
      where(is.numeric),
      list(mean = mean, median = median, sd = sd)
    )
  )
```


## Working with dates
```{r}
hr_data |>
  mutate(
    datetime = lubridate::round_date(collection_datetime, unit = "minutes"),
    .before = collection_datetime
  )
```

## summarising within groups
```{r}
hr_data |>
  mutate(
    collection_datetime = lubridate::round_date(
      collection_datetime,
      unit = "minute"
    )
  ) |>
  summarise(
    across(
      where(is.numeric),
      list(mean = mean, sd = sd, median = median)
    ),
    .by = c(id, collection_datetime)
  )
```

## Summarise as a code
```{r}
#' Summarise to get mean, median, sd
#'
#' @param data dataframe to summarise on
#'
#' @returns summarised data by id and collection_datetime (minutes)

summarise_by_datetime <- function(data, unit = "minutes") {
  summarised_data <- data |>
    dplyr::mutate(
      collection_datetime = lubridate::round_date(
        collection_datetime,
        unit = unit
      )
    ) |>
    dplyr::summarise(
      dplyr::across(
        tidyselect::where(is.numeric),
        list(mean = mean, sd = sd, median = median)
      ),
      .by = c(id, collection_datetime)
    )
  return(summarised_data)
}
```

```{r}
#| eval: false

hr_data <- read_all("HR.csv.gz") |>
  get_participant_id() |>
  summarise_by_datetime()
ibi_data <- read_all("IBI.csv.gz") |>
  get_participant_id() |>
  summarise_by_datetime()
temp_data <- read_all("TEMP.csv.gz") |>
  get_participant_id() |>
  summarise_by_datetime()
bvp_data <- read_all("BVP.csv.gz") |>
  get_participant_id() |>
  summarise_by_datetime()
eda_data <- read_all("EDA.csv.gz") |>
  get_participant_id() |>
  summarise_by_datetime()

# making the above as a function

read_sensor_data <- function(filename) {
  data <- read_all(filename) |>
    get_participant_id() |>
    summarise_by_datetime()
  return(data)
}

# modify to be able to change unit
#' Read_sensor_data
#'
#' @param filename name of file
#' @param unit unit of datetime
#'
#' @returns dataframe that has id and is summarised by datetime

read_sensor_data <- function(filename, unit) {
  data <- read_all(filename) |>
    get_participant_id() |>
    summarise_by_datetime(unit = unit)
  return(data)
}
```


## extra exercise 20.10
```{r}
#| eval: false

# don't run this chunk because the function is different in the functions.R file
hr_data <- read_all("HR.csv.gz") |>
  get_participant_id()

summarise_by_datetime(
  data = hr_data,
  fns = list(median = median),
  unit = "minutes"
)

summarise_by_datetime(
  data = hr_data,
  fns = mean,
  unit = "minutes"
)

summarise_by_datetime(
  data = hr_data,
  fns = list(max = max, min = min),
  unit = "minutes"
)
```

## Pivot longer
```{r}
survey_data <- here("data-raw/nurses-stress/survey-results.csv.gz") |>
  read()
survey_data

survey_data |>
  mutate(
    date = mdy(date),
    start_datetime = as_datetime(
      paste(
        date, start_time
      )
    ),
    end_datetime = as_datetime(
      paste(
        date, end_time
      )
    ),
    .before = start_time
  )

survey_tidy <- survey_data |>
  mutate(
    date = lubridate::mdy(date),
    start_datetime = lubridate::as_datetime(paste(date, start_time)),
    end_datetime = lubridate::as_datetime(paste(date, end_time)),
    datetime_id = start_datetime,
    .before = start_time
  ) |>
  dplyr::select(-c(date, start_time, end_time, duration))

survey_long <- survey_tidy |>
  select(id, datetime_id, start_datetime, end_datetime) |>
  tidyr::pivot_longer(c(start_datetime, end_datetime),
    names_to = NULL,
    values_to = "collection_datetime"
  ) |>
  group_by(dplyr::pick(-collection_datetime)) |>
  tidyr::complete(
    collection_datetime = seq(
      min(collection_datetime),
      max(collection_datetime),
      by = 60
    )
  ) |>
  ungroup()
```

## making the above into functions
```{r}
#| eval: false

#' tidy_survey_dates
#'
#' @param data dataframe survey_data
#'
#' @returns a tidy dataframe
tidy_survey_dates <- function(data) {
  tidied <- data |>
    dplyr::mutate(
      date = lubridate::mdy(date),
      start_datetime = lubridate::as_datetime(paste(date, start_time)),
      end_datetime = lubridate::as_datetime(paste(date, end_time)),
      datetime_id = start_datetime,
      .before = start_time
    ) |>
    dplyr::select(-c(date, start_time, end_time, duration))
  return(tidied)
}

#' survey_to_long
#'
#' @param data tidy dataframe on survey
#'
#' @returns survey data in a longer format dataframe
survey_to_long <- function(data) {
  longer <- data |>
    dplyr::select(id, datetime_id, start_datetime, end_datetime) |>
    tidyr::pivot_longer(c(start_datetime, end_datetime),
      names_to = NULL,
      values_to = "collection_datetime"
    ) |>
    dplyr::group_by(dplyr::pick(-collection_datetime)) |>
    tidyr::complete(
      collection_datetime = seq(
        min(collection_datetime),
        max(collection_datetime),
        by = 60
      )
    ) |>
    dplyr::ungroup()
  return(longer)
}

survey_tidy <- tidy_survey_dates(survey_data)
survey_long <- survey_to_long(survey_tidy)
```



# Notes
Don't use 'for loops' at all -> use purrr package and functionals like map() (applying a function to a vector(multiple objects))

duckdb is much faster than data.table

You can use pipes within mutate (see ## extracting id)
